PineAPPL:\\ combining EW and QCD corrections for PDF fits

### 2020-08-13 paper/sec-introduction.tex
General revision

The two respects are intertwined: PDFs provide a way of obtaining a prediction
for a given process in terms of other processes, therefore the accuracy and 
precision with which PDFs are able to make predictions crucially depend on 
the accuracy of the perturbative computation used in the PDF determination
(matched to the precision of the data). Electroweak (including combined QCD--EW) 
effects should therefore naturally be included in a PDF determination. The 
reason is twofold. First, one expects NNLO QCD and NLO EW corrections to be of 
comparable size because, at the EW scale, the QCD and EW running couplings 
become similar, $\alphas^2\sim \alpha$. If NNLO QCD corrections are included
by default in PDF determinations, NLO EW corrections should be
taken into account as well. Second, the virtual exchange of soft or 
collinear weak bosons leads to Sudakov 
logarithms~\cite{Denner:2000jv,Denner:2001gw},
which can make the coefficients of the EW series grow faster than 
their QCD counterparts. Such a behaviour becomes relevant in
phase-space regions associated with large mass scales (roughly of the order
of the \si{\tera\electronvolt}), where several LHC data sets (e.g.\ Z-boson
transverse-momentum distributions) enter both the determination of PDFs and the search
for new physics.

@@

In this paper we address the first of these two problems. Specifically, 
we develop \textsc{PineAPPL}, a library that allows the user to generate
fast-interpolation grids, accurate to up to any fixed order in the strong and
EW couplings. The library supports variations of the factorisation
and renormalisation scales, and can be extended to include
resummation and matching with a photon- and/or parton-shower.
The grids in the new \textsc{PineAPPL} format complement those
(accurate to fixed-orders in the strong coupling only) that can be generated 
in the \textsc{APPLgrid} and \textsc{FastNLO} formats. The \textsc{PineAPPL}
library is interfaced to \textsc{MadGraph5\_aMC@NLO} (\textsc{MG5\_aMC}
henceforth), with which it has been developed and tested. In this respect,
\textsc{PineAPPL} supersedes
\textsc{APPLgrid}+\textsc{aMCfast}~\cite{Bertone:2014zva}.
\textsc{PineAPPL} can also be easily interfaced with any Monte Carlo generator,
e.g.\ \textsc{SHERPA}~\cite{Biedermann:2017yoi}, where it complements
\textsc{MCgrid}~\cite{DelDebbio:2013kxa,Bothmann:2015dba}.

>>>>

The two respects are intertwined: PDFs provide a way of obtaining a prediction
for a given process in terms of other processes, therefore the accuracy and 
precision with which PDFs are able to make predictions crucially depend on 
the accuracy of the perturbative computation used in the PDF determination
(matched to the precision of the data). Electroweak (including combined QCD--EW) 
effects should therefore naturally be included in a PDF determination. The 
reason is twofold. First, one expects NNLO QCD and NLO EW corrections to be of 
comparable size because, at the EW scale, the QCD and EW running couplings 
become similar, $\alphas^2\sim \alpha$. If NNLO QCD corrections are included
by default in PDF determinations, NLO EW corrections should be
taken into account as well. Second, the virtual exchange of soft or 
collinear weak bosons leads to Sudakov 
logarithms~\cite{Denner:2000jv,Denner:2001gw},
which can make the coefficients of the EW series grow faster than 
their QCD counterparts. This behaviour is relevant in
phase-space regions associated with large mass scales (roughly of the order
of the \si{\tera\electronvolt}), where several LHC data sets (e.g.\ Z-boson
transverse-momentum distributions) enter both the determination of PDFs and the search
for new physics.

@@

In this paper we address the first of these two problems. Specifically, 
we develop \textsc{PineAPPL}, a library that allows any user to generate
fast-interpolation grids, accurate to up to any fixed order in the strong and
EW couplings. The library supports variations of the factorisation
and renormalisation scales, and can be extended to include
resummation, and matching with a photon- and/or parton-shower.
The grids in the new \textsc{PineAPPL} format complement those
(accurate to fixed-orders in the strong coupling only) that can be generated 
in the \textsc{APPLgrid} and \textsc{FastNLO} formats. The \textsc{PineAPPL}
library is interfaced to \textsc{MadGraph5\_aMC@NLO} (\textsc{mg5\_aMC}
henceforth), with which it has been developed and tested. In this respect,
\textsc{PineAPPL} supersedes
\textsc{APPLgrid}+\textsc{aMCfast}~\cite{Bertone:2014zva}.
\textsc{PineAPPL} can also be easily interfaced to any Monte Carlo generator,
e.g.\ \textsc{SHERPA}~\cite{Biedermann:2017yoi}, where it complements
\textsc{MCgrid}~\cite{DelDebbio:2013kxa,Bothmann:2015dba}.
